---
title: "Project1_simulation"
author: "Jinhai_Yu"
date: "11/13/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

The task is to practice the simulation of exponential distribution in R and compare it with the Central Limit Theorem. The report will show the sample mean and variance, and the comparison with the theoretical values, also to see if the distribution is close to normal or not. 

## Load libraries
```{r}
library(ggplot2)
```

## Simulation

The exponential distribution is rexp(n, lambda), with n = 40, lambda = 0.2, mean = 1/0.2, sd = 1/0.2; and the simulation times are 1000.
```{r Simulation}
set.seed(2020)
lambda <- 0.2
n <- 40
rep <- 1000
sim1 <- NULL
sim2 <- NULL
for(i in 1:rep) sim1 <- c(sim1, mean(rexp(n, lambda)))
for(i in 1:rep) sim2 <- c(sim2, var(rexp(n, lambda)))
mean_sim <- mean(sim1)
var_sim <- mean(sim2)
mean_t <- 1/lambda
var_t <- (1/lambda)^2


```
## Plot mean and variance of simulated data, and compared with theoretical values, respectively.

```{r plot histogram}
par(mfrow = c(1, 2))
hist(sim1, breaks = 30, xlab = "Mean", main = " Histogram of mean")
abline(v = mean_sim, col = "blue")
abline(v = mean_t, col = "red")
legend("topright", col = c("blue", "red"), legend = c("Sample mean", "Theoretical mean"), lty = 1, cex = 0.5)

hist(sim2, breaks = 30, xlab = "Variance", main = " Histogram of variance")
abline(v = var_sim, col = "blue")
abline(v = var_t, col = "red")
legend("topright", col = c("blue", "red"), legend = c("Sample variance", "Theoretica variance"), lty = 1, cex = 0.5)
cat(paste(c("The mean of sample mean is ", mean_sim, " that is pretty close to the theoretical mean ", 
            mean_t), sep = "", collapse = ""))
cat(paste(c("The mean of sample var is ", var_sim, " that is also close to the theoretical variance ", 
            var_t), sep = "", collapse = ""))


```

## Plot mean of simulated data with density and compared with normal distribution 
(mean =mean of simulated data, sd = sd of simulated data)
```{r plot density}
sim3 <- as.data.frame(sim1)
ggplot(sim3, aes(x = sim1)) +
        geom_histogram(aes(y = ..density..), binwidth = 0.25, col = "black", fill = "gray") +
        geom_density(aes(color = "Simulated")) +
        geom_function(aes(color = "Normal"), fun = dnorm, args = list(mean = mean_sim, sd = sd(sim1))) +
        labs(x = "Mean", y = "Density", title = "Distribution of simulated data", color ="Legend") +
        theme(plot.title = element_text(hjust = 0.5)) +
        scale_color_manual("Distribution", values = c("red", "blue"))
cat(paste("The density line of simulated data is very similar to the normal distribution."))
```