---
title: "Toothgrowth"
author: "Jinhai_Yu"
date: "11/15/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This project compared the effects of OJ and VC on the tooth growth in guinea pigs,
I did a brief summary with mean and quantile, also a bar plot to show the difference.
At last, I did t test to see if the differences are significant or not with t.test().

## Load library
```{r load_library, echo=FALSE}
library(datasets)
library(dplyr)
library(ggplot2)
```

## Load data and do a brief summary
plot the comparison of tooth lengths between the two supplements according to different dosage

```{r load_data and plot, fig.height=3}
data("ToothGrowth")
dim(ToothGrowth)
str(ToothGrowth)
ToothGrowth %>% group_by(supp) %>% 
        summarise(mean = mean(len), quantile = quantile(len, probs = seq(0, 1, 0.25)))
ggplot(data = ToothGrowth, aes(x = supp, y = len, fill = as.factor(supp))) +
        geom_bar(stat = "identity") +
        facet_grid(. ~ as.factor(dose)) +
        labs(x = "Supplement", y = "Tooth Length", 
             title = "The Effect of Supplements on Tooth Growth in Guinea Pigs") +
        theme(plot.title = element_text(hjust = 0.5)) +
        scale_fill_discrete(name = "Dosage (mg/day)") 
```

## simple t test

```{r t test, tidy=TRUE, tidy.opts=list(width.cutoff=50)}
# test the overall effect between the 2 supplements 
# hypothesis is: there's no difference on the effect of tooth growth between the 2 supplements 
options(width = 60)
test1 <- t.test(len ~ supp, data = ToothGrowth)
test1$conf.int
cat(paste(c("The 95% confidence interval is ", format(test1$conf.int[1], digits = 4), " ", 
            format(test1$conf.int[2], digits = 4), " that includes 0, so there's no significant difference 
            between the two supplements on tooth growth."), sep = "", collapse = ""))
# test the effect between the 2 supplements according to dosage
# hypothesis is: there's no difference on the effect of tooth growth between the 2 supplements with dosage as 0.5, 1, 2
test2 <- t.test(len ~ supp, data = subset(ToothGrowth, dose == 0.5))
test3 <- t.test(len ~ supp, data = subset(ToothGrowth, dose == 1))
test4 <- t.test(len ~ supp, data = subset(ToothGrowth, dose == 2))
cat(paste(c("The p value of t test for the dosage 0.5, 1, 2 are ", 
            format(test2$p.value, digits = 4), " ", format(test3$p.value, digits = 4), " ", format(test4$p.value, digits = 4), 
            " \nrespectively." ," We could see that low dosage of OJ is better thant VC, and the difference 
            is significant; but for high dosage there's no significant difference."), sep = "", collapse = ""))
```

## Conclusion

The overall effect of OJ and VC on tooth growth are not different; however, when considering different dosage, 
OJ showed better effect than VC on tooth growth on lower dosage (0.5 and 1), but there's no difference on high dosage (2). 
